Paper Title*\\ \thanks{Identify applicable funding agency here. If none, delete this.}

### 2019-07-17 paper.tex
Updates from Overleaf

Throughout the fall detection literature, techniques varies from threshold activion based, such as, Sase et al.~\cite{sase2018human} detecting regions of interest after a background subtraction, and a threshold activation function if the region of interest was bellow a third of the subject height in the frame. Bhandari et al.~\cite{bhandari2017novel} determined regions of interest via Shi-Tomasi, tracked them using Lucas-Kanade, and set a threshold on the region of interest's speed and motion. And Albawendi et al.~\cite{albawendi2018video} who used a threshold activation on silhouette shape. Another commonly used classification technique is support vector machine (SVM), used by Harrou et al.~\cite{harrou2017vision} that fed a feature extractor with the subject segmented from the frame. The extractor output would then be compared with a multivariate exponentially-weighted moving average (MEWMA), and classified by a SVM. Panahi et al.~\cite{panahi2018human} subtracted the background from depth images, fitted an ellipse in the subject's shape, and classified it with SVM. Abobakr et al.~\cite{abobakr2017skeleton} also subtracted the background from depth images, applied a random decision forest to estimate the subject's posture, and a SVM to classify it. Regarding the subject's privacy, both Edgcomb et al.~\cite{edgcomb2012automated} and Lin et al.~\cite{lin2013fall} investigated privacy-prone features, such as blurring, silhouetting, and covering the body with shapes.

~\cite{xu2018survey}
~\cite{yang2019.} % saliency
~\cite{sehairi2018elderly} % KNN
In a mixed approach, Kwolek and Kepski~\cite{kwolek2015improving} used an accelerometer, that passed a given threshold would activate a K-nearest neighbors vision classifier to confirm the event. % KNN
~\cite{mastorakis2018.} % simulated db

~\cite{sigal2018interpretable} % salinecy

% markov
Zerrouki et al.~\cite{zerrouki2018combined} that after subtracting the background, applied a curvelet transform and served it to a SVM to classify the human pose, and then a hidden markov model to recognize the activity.

Qian et al.~\cite{qian2017recognizing} also employed a hidden markov model to detect falls.

Yu et al.~\cite{yu2009fall} utilized particle filtering with a background subtraction to track the subject head, and detect falls.

% gaussian
Yu et al.~\cite{yu2010robust} applyed a head tracking based on a gaussian model with a parzen window method to calculate density.

Rougier et al.~\cite{rougier2011robust} detected falls with gaussian and human shape deformation.

~\cite{lu2018deep} % LSTM

~\cite{huang2018video} % CNN

% CNN
N\'u\~nez-Marcos et al.~\cite{nunez2017vision} extracted dense optical flow (OF) from images and fed them to a VGG-16 convolutional neural network (CNN). The extracted optical flow was fed in stacks to create a temporal relation between frames.

Anishchenko et al.~\cite{anishchenko2018machine} employd a modified AlexNet to detect falls.

\section{Background}

In this section, we describe a few relevant concepts applied in the proposed method.

\subsection{Optical Flow}
\label{sec:opticalflow}

Optical flow is a hand-crafted feature that conveys movement information between the frames of a video, either by object movement or camera shift, Figure~\ref{fig:opticalflow}. Its ability to describe movement enables this feature to support the deep network towards detecting falls.

The extraction process considers a video frame as $I$ and $I(x, y, t)$ a pixel in this frame. Another frame collected $dt$ time later, and the distance that the same pixel moved $(dx, dy)$, described in Equation~\ref{eq:of-dist}.

Optical flow Equation~\ref{eq:of} is formed after applying a Taylor series approximation of right-hand side and dividing by $dt$, $f_t$ is the gradient given time, $f_x$, $f_y$, $u$, and $v$ are given in Equation~\ref{eq:of-grad}. Values of $u$ and $v$ can be achieved by different methods, such as the classical Lucas-Kanade~\cite{jain2018abnormal} or the one used in this work: Gunnar-Farneb{\"a}ck~\cite{lowhur2015dense}.

\begin{equation}
\label{eq:of-dist}
I(x, y, t)=I(x+dx, y+dy, t+dt)
\end{equation}
\begin{equation}
\label{eq:of}
f_xu + f_yv + f_t=0
\end{equation}
\begin{equation}
\label{eq:of-grad}
f_x = \frac{\partial f}{\partial x}; \quad f_y = \frac{\partial f}{\partial y}u = \frac{\partial x}{\partial t}; \quad v = \frac{\partial y}{\partial t}
\end{equation}

\subsection{Saliency Map}
\label{sec:saliency}

Saliency was envisioned as a visualizing technique to understand the learning process behind a deep network. It can be extracted in many ways, one of those is the practice of feeding a image $I$ to the network and its class label $c$, freeze the weights on the network and compute the gradients of $c$ with respect to the pixels of $I$. Consider $S_c(I)$ to be the class score function for an image $I$ and, for this simple explanation, that it is a simple linear function, $S_c(I) = w_c^{T} + b_c$. With $w$ being the weights of the network and $b$ the bias, it is possible to see that the magnitude of elements of $w$ affects the importance of the corresponding pixels in $I$ for class $c$.

The saliency map was chosen as a stream for its spatial relation to the frame, as the raw RGB frames are prone to interference, for instance from brightness variation, we believe the saliency is capable to overcome these variables. This work's saliency extractor is implemented after the paper of Sundararajan et al.~\cite{sundararajan2017axiomatic}.

\section{Method}

@@

Two selected features were fed to the streams explained in Subsection~\ref{sec:classification}:(i) Optical Flow and (ii) Saliency Map. Furthermore, the RGB frame itself was fed to the third stream, Figure~\ref{fig:features}, given the neural networks capability to learn relevant features to describe their input.

\paragraph{Optical Flow} represents the temporal relation between frames, and to fully gather this property it was used as a stack of optical flows on top of a sliding window of frames. The sliding window has a size of $L$ frames and gathers 2$L$ components ($L$ horizontal ($d_t^x$) + $L$ vertical ($d_t^y$) optical flow component vector fields) creating a stack $O$ = \{$d_t^x$, $d_t^y$, $d_{t+1}^x$, $d_{t+1}^y$, ..., $d_{t+L}^x$, $d_{t+L}^y$ \}. The total number of stacks is given by $N$-$L$+$1$, where $N$ is the number of frames in a video, Figure~\ref{fig:stacks, fig:features}. 
\paragraph{Saliency Map} is composed of pixels varying from 0 to 1, and was fed as a gray-scale image to the network, Figure~\ref{fig:features}.

@@

\begin{table*}[t]
\centering
\caption{URFD single-stream \textit{vs} multi-stream, decreasing accuracy.}
\label{tab:urfd-ensem}
\begin{tabular}{llcccl}
\hline
 &                              & Sensitivity (\%)  & Specificity (\%)  & Accuracy (\%)     & \\ \hline
 & Multi-stream (OF+RGB+Sal)    & \textbf{100}      & \textbf{98.77}    & \textbf{98.84}    & \\
 & Single-stream (OF)           & \textbf{100}      & 97.42             & 97.57             & \\
 & Single-stream (RGB)          & \textbf{100}      & 96.35             & 96.56             & \\
 & Single-stream (Sal)          & 93.67             & 93.39             & 93.40             & \\ \hline
\end{tabular}
\end{table*}

@@

\begin{table*}[t]
\centering
\caption{FDD single-stream \textit{vs} multi-stream, decreasing accuracy.}
\label{tab:fdd-ensem}
\begin{tabular}{llcccl}
\hline
 &                      & Sensitivity (\%)  & Specificity (\%)  & Accuracy (\%)     &  \\ \hline
 & Multi (OF+RGB+Sal)   & \textbf{99.43}    & \textbf{99.55}    & \textbf{99.51}    &  \\
 & Single (OF)          & 99.07             & 99.12             & 99.11             &  \\
 & Single (RGB)         & 98.85             & 98.41             & 98.55             &  \\
 & Single (Sal)         & 88.39             & 92.40             & 91.25             & \\ \hline
\end{tabular}
\end{table*}

@@

\section{Conclusion and Future Work}

>>>>

Throughout the fall detection literature, techniques varies from threshold activation based, such as, Sase et al.~\cite{sase2018human} detecting regions of interest after a background subtraction, and a threshold activation function if the region of interest was bellow a third of the subject height in the frame. Bhandari et al.~\cite{bhandari2017novel} determined regions of interest via Shi-Tomasi, tracked them using Lucas-Kanade, and set a threshold on the region of interest's speed and motion. And Albawendi et al.~\cite{albawendi2018video} who used a threshold activation on silhouette shape. Another commonly used classification technique is support vector machine (SVM), used by Harrou et al.~\cite{harrou2017vision} that fed a feature extractor with the subject segmented from the frame. The extractor output would then be compared with a multivariate exponentially-weighted moving average (MEWMA), and classified by a SVM. Panahi et al.~\cite{panahi2018human} subtracted the background from depth images, fitted an ellipse in the subject's shape, and classified it with SVM. Abobakr et al.~\cite{abobakr2017skeleton} also subtracted the background from depth images, applied a random decision forest to estimate the subject's posture, and a SVM to classify it. Regarding the subject's privacy, both Edgcomb et al.~\cite{edgcomb2012automated} and Lin et al.~\cite{lin2013fall} investigated privacy-prone features, such as blurring, silhouetting, and covering the body with shapes. Deep neural approaches were explored by Lu et al.~\cite{lu2018deep} who employed a 3D convolutional neural network (CNN) followed by a long short-term memory (LSTM). Huang et al.~\cite{huang2018video} extracted the human pose with OpenPose and classified it with a VGG-16 architecture. N\'u\~nez-Marcos et al.~\cite{nunez2017vision} extracted the dense optical flow (OF) and also served it to a VGG-16. And Anishchenko et al.~\cite{anishchenko2018machine} employed a modified AlexNet as classifier. A few different approaches were also reported, such as, K-nearest neighbor classifiers by Sehairi et al.~\cite{sehairi2018elderly} and Kwolek et al.~\cite{kwolek2015improving}. Hidden markov model by Zerrouki et al.~\cite{zerrouki2018combined} and Qian et al.~\cite{qian2017recognizing}. Particle filtering by Yu et al.~\cite{yu2009fall}. And gaussian models by Yu et al.~\cite{yu2010robust} and Rougier et al.~\cite{rougier2011robust}. Furthermore Xu et al.~\cite{xu2018new} produced a survey on fall detections systems.

This work's main contribution is the development and evaluation of hand-crafted features to be used in a multi-stream learning model. To this end, we compare the effectiveness of three features extracted from videos frames: (i) optical flow, (ii) saliency map, and (iii) RGB features. These features a fed to a three-stream VGG-16 architecture, and their results compared with the literature.

This paper is organized as follows. In Section~\ref{sec:method} we detail the proposed three-stream methodology. In Section~\ref{sec:experiments} the performed experiments and their results are compared to other works in the literature. Moreover, in Section~\ref{sec:conclusion} we elaborate on some points and future work suggestions.

% ~\cite{yang2019.} % saliency
% ~\cite{mastorakis2018.} % simulated db

% ~\cite{sigal2018interpretable} % salinecy

% \section{Background}
% \label{background}

% In this section, we describe a few relevant concepts applied in the proposed method.

% \subsection{Optical Flow}
% \label{sec:opticalflow}

% Optical flow is a hand-crafted feature that conveys movement information between the frames of a video, either by object movement or camera shift, Figure~\ref{fig:opticalflow}. Its ability to describe movement enables this feature to support the deep network towards detecting falls.

% The extraction process considers a video frame as $I$ and $I(x, y, t)$ a pixel in this frame. Another frame collected $dt$ time later, and the distance that the same pixel moved $(dx, dy)$, described in Equation~\ref{eq:of-dist}.

% Optical flow Equation~\ref{eq:of} is formed after applying a Taylor series approximation of right-hand side and dividing by $dt$, $f_t$ is the gradient given time, $f_x$, $f_y$, $u$, and $v$ are given in Equation~\ref{eq:of-grad}. Values of $u$ and $v$ can be achieved by different methods, such as the classical Lucas-Kanade~\cite{jain2018abnormal} or the one used in this work: Gunnar-Farneb{\"a}ck~\cite{lowhur2015dense}.

% \begin{equation}
% \label{eq:of-dist}
% I(x, y, t)=I(x+dx, y+dy, t+dt)
% \end{equation}
% \begin{equation}
% \label{eq:of}
% f_xu + f_yv + f_t=0
% \end{equation}
% \begin{equation}
% \label{eq:of-grad}
% f_x = \frac{\partial f}{\partial x}; \quad f_y = \frac{\partial f}{\partial y}u = \frac{\partial x}{\partial t}; \quad v = \frac{\partial y}{\partial t}
% \end{equation}

% \subsection{Saliency Map}
% \label{sec:saliency}

% Saliency was envisioned as a visualizing technique to understand the learning process behind a deep network. It can be extracted in many ways, one of those is the practice of feeding a image $I$ to the network and its class label $c$, freeze the weights on the network and compute the gradients of $c$ with respect to the pixels of $I$. Consider $S_c(I)$ to be the class score function for an image $I$ and, for this simple explanation, that it is a simple linear function, $S_c(I) = w_c^{T} + b_c$. With $w$ being the weights of the network and $b$ the bias, it is possible to see that the magnitude of elements of $w$ affects the importance of the corresponding pixels in $I$ for class $c$.

% The saliency map was chosen as a stream for its spatial relation to the frame, as the raw RGB frames are prone to interference, for instance from brightness variation, we believe the saliency is capable to overcome these variables. This work's saliency extractor is implemented after the paper of Sundararajan et al.~\cite{sundararajan2017axiomatic}.

\section{Method}
\label{sec:method}

@@

Two selected features were fed to the streams explained in Subsection~\ref{sec:classification}:(i) Optical Flow and (ii) Saliency Map. Furthermore, the RGB frame itself was fed to the third stream, given the neural networks capability to learn relevant features to describe their input.

\paragraph{Optical Flow} represents the temporal relation between frames, and to fully gather this property it was used as a stack of optical flows on top of a sliding window of frames. The sliding window has a size of $L$ frames and gathers 2$L$ components ($L$ horizontal ($d_t^x$) + $L$ vertical ($d_t^y$) optical flow component vector fields) creating a stack $O$ = \{$d_t^x$, $d_t^y$, $d_{t+1}^x$, $d_{t+1}^y$, ..., $d_{t+L}^x$, $d_{t+L}^y$ \}. The total number of stacks is given by $N$-$L$+$1$, where $N$ is the number of frames in a video, Figure~\ref{fig:stacks}. 
\paragraph{Saliency Map} is composed of pixels varying from 0 to 1, and was fed as a gray-scale image to the network, Figure~\ref{fig:features}.

@@

\begin{table}[]
\centering
\caption{URFD single-stream \textit{vs} multi-stream, decreasing accuracy.}
\label{tab:urfd-ensem}
\begin{tabular}{llcccl}
\hline
 &  & Sens. (\%) & Spec. (\%) & Acc. (\%) &  \\ \hline
 & OF + RGB + Sal & \textbf{100} & \textbf{98.77} & \textbf{98.84} &  \\
 & Optical Flow & \textbf{100} & 97.42 & 97.57 &  \\
 & RGB & \textbf{100} & 96.35 & 96.56 &  \\
 & Saliency & 93.67 & 93.39 & 93.40 &  \\ \hline
\end{tabular}
\end{table}

@@

\begin{table}[]
\centering
\caption{FDD single-stream \textit{vs} multi-stream, decreasing accuracy.}
\label{tab:fdd-ensem}
\begin{tabular}{llcccl}
\hline
 &  & Sens. (\%) & Spec. (\%) & Acc. (\%) &  \\ \hline
 & OF + RGB + Sal & \textbf{99.43} & \textbf{99.55} & \textbf{99.51} &  \\
 & Optical Flow & 99.07             & 99.12             & 99.11             &  \\
 & RGB & 98.85             & 98.41             & 98.55             &  \\
 & Saliency & 88.39             & 92.40             & 91.25             & \\ \hline
\end{tabular}
\end{table}

@@

\section{Conclusion and Future Work}
\label{sec:conclusion}
